---
title: "Underground vs surface by country"

---

Looking at June 2018 Mining Intelligence data.

 

```{r}
library(tidyverse)

rawpropaddress <- read_csv("C:/Users/mmccrae/Desktop/property addresses format r june 18 2018.csv")
View(rawpropaddress)

rawprocess <- read_csv("C:/Users/mmccrae/Desktop/processing r format june 18 2018.csv")
View(rawprocess)

overviewraw <- read_csv("C:/Users/mmccrae/Desktop/overview r format june 2018.csv")
View(overviewraw)

glimpse(rawpropaddress)
glimpse(rawprocess)
glimpse(overviewraw)


```



K. Let's join two data sets. Just join overviewraw and rawprocess since the other dataset appears to be broken. 



```{r}

joinraw <- full_join(rawprocess,overviewraw)
View(joinraw)
```


What country has the most operating mines? 

```{r}

mostmines <- joinraw %>% filter(AssetType=="Mine") %>% filter(ActivityStatus=="Production") %>% filter(WorkType %in% c("Open-Pit", "Underground")) %>% group_by(Country) %>% tally(sort=TRUE) %>% top_n(20)
mostmines
```



Let's chart. 


```{r}



library(scales)
library(showtext)
library(extrafont)


mm <- ggplot(mostmines, aes(x=reorder(Country,n), y=n))+
  geom_bar(stat="identity",width=.1, fill="#918f8f")+
  geom_point(size=6, color="#918f8f")+
  geom_text(aes(label=n), hjust = -.9, family="Montserrat")+
  coord_flip() +
  scale_y_continuous(limits = c(0, 1000), labels = scales::comma) +
  theme_minimal(base_size=16, base_family="Montserrat")+
  labs(title="Number of Mines in Production by Country", subtitle="Mines classified as underground or open pit as of June 2018. Showing just the top 20 countries.",x="", y="", caption="Mining Intelligence", fill="") +
  theme(plot.title = element_text(size = rel(2), family="Zilla Slab Light"), legend.title=element_blank()) +
  scale_colour_manual(values=c("#99d594","#fc8d59"))
#  theme(legend.position="none")

mm


  ggsave("minesbycountry4.pdf", plot=mm, width=280, height=210, units="mm")
  
```





Let's look at top 20 country with underground and surface. 




```{r}

#undersurf <- joinraw %>% filter(Country %in% c("United States","Russia","Australia","China","India","South Africa","Canada","Ukraine","Brazil","Mexico","Indonesia","Peru","Kazakhstan","Chile","Zimbabwe","Poland","Vietnam","Turkey","Dem. Republic of the Congo","Philippines")) %>% filter(ActivityStatus=="Production") %>% filter(WorkType %in% c("Open-Pit", "Underground")) %>% group_by(Country, WorkType) %>% tally(sort=TRUE) %>% ungroup() %>% group_by(Country) %>% tally() %>% ungroup()

undersurf <- joinraw %>% filter(Country %in% c("United States","Russia","Australia","India","China","South Africa","Ukraine","Canada","Brazil","Mexico","Indonesia","Peru","Chile","Kazakhstan","Zimbabwe","Poland","Vietnam","Philippines", "Turkey", "Germany")) %>%  filter(AssetType=="Mine") %>% filter(ActivityStatus=="Production") %>% filter(WorkType %in% c("Open-Pit", "Underground")) %>% group_by(Country, WorkType) %>% tally(sort=TRUE)





undersurf$Country <- factor(undersurf$Country, levels=c("United States","Russia","Australia","India","China","South Africa","Ukraine","Canada","Brazil","Mexico","Indonesia","Peru","Chile","Kazakhstan","Zimbabwe","Poland","Vietnam","Philippines", "Turkey", "Germany"))

undersurf$Country <- fct_rev(undersurf$Country)

glimpse(undersurf)
undersurf

```



Let' spread to calculate values of mean


```{r}


spreadundersurf <- undersurf %>% spread(WorkType,n) %>% mutate(totalprodmines = `Open-Pit`+Underground) %>% mutate(OpenPitPercent=`Open-Pit`/totalprodmines) %>% mutate(UndergroundPercent=Underground/totalprodmines)
spreadundersurf 


```


Now time to gather to chart. 


```{r}

gatherundersurf <- spreadundersurf %>% gather(key=WorkTypeName, value=WorkTypeValue, OpenPitPercent,UndergroundPercent)
gatherundersurf

```






Let's plot. 

```{r}

ggplot(gatherundersurf, aes(x=Country, WorkTypeValue, fill=WorkTypeName))+
  geom_bar(stat="identity")+
  coord_flip()

```







Let's format to make prettier. 

```{r}

library(scales)
library(showtext)
library(extrafont)

ee <- ggplot(gatherundersurf, aes(x=Country, WorkTypeValue, fill=WorkTypeName))+
  geom_bar(stat="identity")+
  coord_flip()+
#  geom_text(aes(label=scales::percent(WorkTypeValue)), family="Montserrat")+
  scale_y_continuous(labels = scales::percent) +
  theme_minimal(base_size=16, base_family="Montserrat")+
  labs(title="Underground and Surface Mines Percentage", subtitle="Showing just the top 20 mining countries as of June 2018.",x="", y="", caption="Mining Intelligence", fill="") +
  theme(plot.title = element_text(size = rel(2), family="Zilla Slab Light"), legend.title=element_blank()) +
  scale_fill_manual(values=c("#e9820d","#2d70da"), labels = c("Open Pit", "Underground"))

ee

  ggsave("pctsurfaceunderground8.pdf", plot=ee, width=280, height=210, units="mm")

```



Let's do some analysis. Count worktype for mines. 

```{r}

undervssurf <- joinraw %>% filter(AssetType=="Mine") %>% filter(ActivityStatus=="Production")%>% group_by(WorkType) %>% tally(sort=TRUE)
undervssurf
```







```{r}

moreanalysis <- joinraw %>% filter(AssetType=="Mine") %>% filter(ActivityStatus=="Production")%>%filter(WorkType %in% c("Open-Pit", "Underground")) %>% group_by(WorkType) %>% summarize(percentage=mean(WorkType))
moreanalysis
```




